---
title: "STAT/MATH 495: Problem Set 07"
author: "Harrison Marick"
date: "2017-10-24"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    collapsed: false
    smooth_scroll: false
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, fig.width=8, fig.height=4.5, message=FALSE, warning = FALSE
  )
set.seed(76)

# Load packages
library(tidyverse)
library(broom)
library(knitr)

train <- read_csv("data/cs-training.csv") %>% 
  rename(Id = X1)
test <- read_csv("data/cs-test.csv") %>% 
  rename(Id = X1)
submission <- read_csv("data/sampleEntry.csv")
```

Information on the competition can be found [here](https://www.kaggle.com/c/GiveMeSomeCredit/data).



# Collaboration

Please indicate who you collaborated with on this assignment: 

I ride solo.


# Build binary classifier

Build the binary classifier based on a single predictor variable: `DebtRatio`,
`age`, or `MonthlyIncome`. Justify this choice.

```{r}
colnames(train)[c(5,11)]=c("late1", "late2") #renaming the NumberOfTime30-59Days... column
colnames(test)[c(5,11)]=c("late1", "late2") #renaming the NumberOfTime60-89Days... column
```


In my exploration of this dataset, I examined various boxplots that indicated the relationship between quantitative variables and our classifier response variable. Below is the boxplot for age, broken down by delinquency status. 

```{r}
ggplot(train, aes(as.factor(SeriousDlqin2yrs), age)) + geom_boxplot() + 
  xlab("Serious Delinquency in Last 2 Years") +
  ylab("Age")
```

Notice the mean age is much lower for the delinquency group, which makes sense as older people tend to be more financial well-off and more responsible than younger people. While this is a generalization, the data suggests there is truth to my intuition.

Before selecting this variable as my predictor of choice, I further investigated by looking at the rate of delinquency by the Age. 

```{r}
proportions <- train %>% 
  group_by(SeriousDlqin2yrs, age) %>% 
  summarise(n=n()) %>% 
  # Create new grouping structure to compute proportions:
  group_by(age) %>% 
  mutate(prop = n/sum(n)) %>% 
  filter(SeriousDlqin2yrs==1)

ggplot(proportions, aes(x=age, y=prop)) +
  geom_line() +
  ylab("Proportion of Serious Delinquency in Last 2 Years") +
  xlab("Age")

```

As we increase the X value, however, the rate of delinquency decreases. This is another indication that age is a good predictor of delinquency. Below, we have fit a logistic model with age as our predictor.

```{r}
mod<-glm(SeriousDlqin2yrs~age, data=train, family="binomial")
log_odds_hat <- predict(mod, newdata=test)
p_hat <- 1/(1 + exp(-log_odds_hat))

test$Probability=p_hat
submission<-test[,c(1,13)] #selecting specific columns
write.csv(submission, "submission.csv", row.names = FALSE)
```



# ROC curve

Based on the ultimate classifier you choose, plot a corresponding ROC curve.

```{r}
train_aug <- mod %>% 
  broom::augment() %>% 
  as_tibble() %>% 
  mutate(p_hat = 1/(1+exp(-.fitted)))

library(ROCR)
# This bit of code computes the ROC curve
pred <- prediction(predictions = train_aug$p_hat, labels = train_aug$SeriousDlqin2yrs)
perf <- performance(pred, "tpr","fpr")

# This bit of code computes the Area Under the Curve
auc <- as.numeric(performance(pred,"auc")@y.values)
auc

plot(perf, main=paste("Area Under the Curve =", round(auc, 3)))

```



# ROC curve for random guessing

Instead of using any predictor information as you did above, switch your
predictions to random guesses and plot the resulting ROC curve.

```{r}
p_hat_random<-runif(length(train_aug$SeriousDlqin2yrs)) #draw from Unif(0,1) dist
pred <- prediction(predictions = p_hat_random, labels = train_aug$SeriousDlqin2yrs)
perf <- performance(pred, "tpr","fpr")

# This bit of code computes the Area Under the Curve
auc <- as.numeric(performance(pred,"auc")@y.values)
auc

plot(perf, main=paste("Area Under the Curve =", round(auc, 3)))
```
